Capstone Movielens Report

### 2025-05-07 reports/capstone-movielens-report.draft4.Rmd
RMD report updated.

rmse2 <- function(true_ratings, predicted_ratings) {
  rmse(true_ratings - predicted_ratings)
}

@@

If we predict all unknown ratings with $\hat{\mu}$, we obtain the following RMSE:

```{r eval=FALSE}
  mu.RMSE <- sqrt(mean(MSEs))
```
```{r}
mu.RMSE
```

If we plug in any other number, we will get a higher RMSE. Let's prove that by the following small investigation:
```{r eval=FALSE}
deviation <- seq(0, 6, 0.1) - 3
print(deviation)

rmse_value <- sapply(deviation, function(diff){
  rmse(test_set$rating - mu + diff)
})

plot(deviation, rmse_value, type = "l")

sprintf("Minimum RMSE is achieved when the deviation from the mean is: %s", 
        deviation[which.min(rmse_value)])

>>>>

rmse2 <- function(true_ratings, predicted_ratings) {
  rmse(true_ratings - predicted_ratings)
}
```

@@

If we predict all unknown ratings with $\hat{\mu}$, we obtain the following RMSE: 
```{r eval=FALSE}
mu.MSEs <- naive_model_MSEs(mu)
```
```{r}
  data.frame(fold_No = 1:5, MSE = mu.MSEs) |>
    data.plot(title = "MSE resuls of the 5-fold CV method applied to the Overall Mean Rating Model",
                xname = "fold_No", 
                yname = "MSE")

mu.RMSE <- sqrt(mean(mu.MSEs))
mu.RMSE
```
::: {.noteblock data-latex=""}
For the _Mean Squared Error_ data visualization we used [`data.plot` function](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R#L630) defined in [_Model Data Visualization_ section](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R#L484) of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R) script.
:::

Here we also used [`naive_model_MSEs` function](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R#L54) defined in the [same script](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R) to compute _Mean Squared Errors_ using _5-Fold Cross Validation_ method:

```{r eval=FALSE}
naive_model_MSEs <- function(val) {
  sapply(edx_CV, function(cv_item){
    mse(cv_item$validation_set$rating - val)
  })
}
```

One more function we will need for further analysis of the current model is the [`naive_model_RMSE` one](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R#L59) defined in the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R) script already mentioned above:
```{r}
naive_model_RMSE <- function(val){
  sqrt(mean(naive_model_MSEs(val)))
}
```

#### Ensure that `mu.RMSE` value is the best for the current model
\

If we plug in any other number, we will get a higher RMSE. Let's prove that by the following small investigation:
```{r eval=FALSE}
  deviation <- seq(0, 6, 0.1) - 3

  deviation.RMSE <- sapply(deviation, function(delta){
    naive_model_RMSE(mu + delta)
  })
  put_log1("RMSE values have been computed for %1 deviations from the Overall Mean Rating.",
           length(deviation))
```

Let's make a quick investigation of the `deviation.RMSE` result we have just got: 
```{r}
data.frame(delta = deviation, 
           delta.RMSE = deviation.RMSE) |> 
data.plot(title = TeX(r'[RMSE as a function of deviation ($\delta$) from the Overall Mean Rating ($\hat{mu}$)]'),
              xname = "delta", 
              yname = "delta.RMSE", 
              xlabel = TeX(r'[$\delta$]'), 
              ylabel = "RMSE")

which_min_deviation <- deviation[which.min(deviation.RMSE)]
min_rmse = min(deviation.RMSE)

put_log1("Minimum RMSE is achieved when the deviation from the mean is: %1",
         which_min_deviation)

put_log1("Is the previously computed RMSE the best for the current model: %1",
         mu.RMSE == min_rmse)
